### 2024-03-12 outputs/paper/paper.qmd
update data and paper

---
title: "Labour Market Resilience"
subtitle: "Economic Crisis Has Diverse Impacts on Industry Work Hours"
author: 
  - Mingjia Chen
  - Catherine Punnoose
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Labour_Market_Resilience.git"
date: March 11, 2024
date-format: long
abstract: "First sentence, second sentence, third sentence."
format: pdf
toc: true
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(janitor,here,dplyr,tidyverse,knitr,kableExtra,ggplot2)

library(janitor)
library(tidyverse)
library(knitr)
library(dplyr)
library(kableExtra)
library(here)
library(ggplot2)
```

# Introduction

# Data

## Source Data

For the analysis, we retrieved the following data as described in @tbl-use-var.

| Variable | New Name   | Description                     | Example |
|----------|------------|---------------------------------|---------|
| response | work hours | Respondent's response           | 23      |
| year     | year       | The year of the survey recorded | 2007    |

: Source data retrieved from GSS {#tbl-use-var}

## Data Cleaning

The data was downloaded and filtered for the selected variables from the selected data variables from GSS[^1]. The data cleaning was performed based on value definitions as defined in the GSS codebook [@gssCodebook]. One of the variable names `response` is renamed to be more informative [@tbl-use-var].

The data was thus been cleaned by using the open source statistically programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], and `knitr` [@knitr].

[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip

```{r}
#| label: tbl-data-overview
#| tbl-cap: Overview of the Categorized Data 
#| echo: false
#| warning: false
#| message: false

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
# Delete the column
data <- data[c(2:ncol(data))]
colnames(data)[1] <- "Work Hours/Years"
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))

data_selected <- data[c("Work Hours/Years", "1998","2008","2021","Total")]
kable(data_selected,row.names = FALSE)|>
  kable_styling() |>
  row_spec(6, hline_after = TRUE)
```

## Data Limitations

## Survey Methodology

### 1998 Survey Methodology 
```{r}
#| label: fig-Survey-1998
#| tbl-cap: Overview of the 1998 Survey Data 
#| echo: false
#| warning: false
#| message: false
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X1998, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "1998 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```
\break

### 2008 Survey Methodology

```{r}
#| label: fig-Survey-2008
#| tbl-cap: Overview of the 2008 Survey Data 
#| echo: false
#| warning: false
#| message: false

data <-read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X2008, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "2008 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```
\break

### Gap in 2020 Data 
```{r}
#| label: fig-Survey-2021
#| tbl-cap: Overview of the 2021 Survey Data 
#| echo: false
#| warning: false
#| message: false
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X2021, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "2021 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```

### Non-responses

# Results

## Working Hour Comparison Between Financial Events

```{r}
#| label: fig-ave-hour-comparison
#| fig-cap: Average Working Hours Comparison between Financial Events
#| echo: false
#| warning: false
#| message: false

data <- read.csv(here::here("inputs/data/Cleaned_Data.csv"))
# Get rid of the first letter X for each of the years
colnames(data)[1] <- "work_hours"
colnames(data) <- gsub("X", "", colnames(data))

#Histogram of average working hours for 2008, 2021, 2022, and Total
hist_data <- select(data, work_hours,"2010","2018","2021")
hist_data <- hist_data[5:94,]

num <- 3 #number of years
sum <- rep(0,num)
total_people <- rep(0,num)
for(i in 1:num){
  total_people[i] <- sum(hist_data[, i+1])
}
hist_data[, 1] <- sapply(hist_data[, 1], as.numeric)
hist_data[90,1] <- 90

sum[1] <- sum(hist_data$"2010" * hist_data$work_hours)
sum[2] <- sum(hist_data$"2018" * hist_data$work_hours)
sum[3] <- sum(hist_data$"2021" * hist_data$work_hours)

averages <- round(sum/total_people) 
years <- c("2010","2018","2021")

average_hours <- data.frame(cbind(averages,years))

ggplot(average_hours,aes(x=years,y=averages)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Year", y = "Average Working Hour") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

\break

```{r}
#| label: fig-cate-comparisons
#| fig-cap: Categorical Working Hour Comparison between the Years
#| fig-subcap: ["Asian Financial Crisis (1998)","World Financial Crisis(2008)","During Covid-19 Pandemic (2021)", "Total"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

#still need to modify the graphs

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))

Data_1998 <- select(data,work_hours,`1998`)
ggplot(Data_1998,aes(x=work_hours,y=`1998`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2008 <- select(data,work_hours,`2008`)
ggplot(Data_2008,aes(x=work_hours,y=`2008`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2021 <- select(data,work_hours,`2021`)
ggplot(Data_2021,aes(x=work_hours,y=`2021`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_Total <- select(data,work_hours,Total)
ggplot(Data_Total,aes(x=work_hours,y= Total)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

\break

## Nonreponse Rate
```{r}
#| label: fig-nonresponse-rate
#| fig-cap: Nonresponse Rates
#| echo: false
#| warning: false
#| message: false

# Calculate percentages
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
data <- data[1:6,]

#colnames(data) <- paste(colnames(data),"Year",sep="_")

colnames(data)<- as.character(colnames(data))

#data_long <- data %>% 
#  pivot_longer(cols = everything(), names_to = "WorkHours", values_to #= "Response")

# Select the 'No Response' row and remove the first column which is not needed for the plot
no_response_data <- data[data$work_hours == 'No Response', -1]

# We also need to remove the 'Total' column as we did in the Python example
no_response_data <- no_response_data[-length(no_response_data)]

# Calculate total responses per year excluding the 'work_hours' and 'Total' columns

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(3:ncol(data))]

total_responses_per_year <- as.vector(data[7,])
total_responses_per_year <- unlist(total_responses_per_year)

# Calculate non-response rate as a percentage
non_response_rate <- (no_response_data / total_responses_per_year) * 100

years <- colnames(non_response_rate)
years <- gsub("X", "", years)

non_response_rate <-as.vector(non_response_rate)
non_response_rate <-unlist(non_response_rate)
non_response_data <- data.frame(years,non_response_rate)

# Plot using ggplot2
ggplot(non_response_data, aes(x = years,y=non_response_rate)) +
  geom_bar(stat = "identity",fill = "skyblue") +
  theme_minimal() +
  labs(title = "Non-Response Rate by Year", x = "Year", y = "Percentage of Non-Responses") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

\break

# Discussion

## Labour Market Response Across Economic Crises

## Demographic Disparities in Labour Market Participation

### Younger Workers and Educational Attainment

### Women in the Labour Market

## Political Affiliation and Labour Market Perceptions

##  Ethical Considerations and Bias

## Limitations

## Future Research

## Journal Study: Directions on Non-response Rates

# Conclusion

\appendix

# Appendix {-}

\newpage
# References